---
title: "Results EC50 assay for MTZ August 2019"
author: "Sascha Krakovka"
#date: "`r Sys.Date()`"
output: 
html_notebook:
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: true
    df_print: paged
    code_folding: hide
    fig_width: 6
    fig_height: 6
---


```{r , include=FALSE}
#Call for the needed libraries in order of usage
library(tidyverse)
library(DT)
library(modelr)
options(na.action = na.warn)
library(knitr)
opts_knit$set(progress = FALSE, verbose = FALSE)
opts_chunk$set(warning=FALSE, message=FALSE, echo=FALSE)
library(tinytex)
library(grid)
library(ggplot2)
```



#Results
###Take all survival rates for the four strains under one drug and combine them to calculate a EC50
Here I take all survival rates from the diverent biological replicates and fit a curve as smooth curve. 

```{r}
a <- read_csv("MTZ_1.csv")
b <- read_csv("MTZ_2.csv")
c <- read_csv("MTZ_3.csv")
d <- read_csv("MTZ_4.csv")
e <- read_csv("MTZ_5.csv")
f <- read_csv("MTZ_6.csv")
MTZ <- rbind(a,b,c,d,e,f)
#MTZ <- rbind(e,f)



#Filter to only keep values that actually show the drug effect and not the noice. I choose 90% as upper limit and 5% as lower limit
MTZ_WB <- filter(MTZ, WB<=105 & WB >= 0.1)
MTZ_C6NR <- filter(MTZ, C6NR<=105 & C6NR >= 0.1)
MTZ_C6aI <- filter(MTZ, C6aI<=105 & C6aI >= 0.1)
MTZ_C60.5 <- filter(MTZ, C60.5<=105 & C60.5 >=0.1)

```


```{r}
#See how it looks and think about y scale for this one
# Plot it 
ggplot()+
    labs(title ="IC50 MTZ")+
    ylab("Procent survival MTZ")+
    xlab("Concentration metronidazole in µM")+
    theme(plot.caption = element_text(size = 11))+
    geom_point (mapping = aes(x = Conc, y = WB), data = MTZ_WB, color="red")+
    geom_point (mapping = aes(x = Conc, y = C6NR), data = MTZ_C6NR, color="blue")+
    geom_point (mapping = aes(x = Conc, y = C60.5), data = MTZ_C60.5, color="green")+
    geom_point (mapping = aes(x = Conc, y = C6aI), data = MTZ_C6aI, color="black")+ 
    stat_smooth(mapping = aes(x = Conc, y = WB), data = MTZ_WB, color="red", formula = y ~log(x))+
    stat_smooth(mapping = aes(x = Conc, y = C6NR), data = MTZ_C6NR, color="blue", formula = y ~log(x))+
    stat_smooth(mapping = aes(x = Conc, y = C60.5), data = MTZ_C60.5, color="green", formula = y ~log(x))+
    stat_smooth(mapping = aes(x = Conc, y = C6aI), data = MTZ_C6aI, color="black", formula = y ~log(x))+
    theme_bw() +
    scale_y_continuous(limits = c(0.1,105))+
    #scale_y_continuous(limits = c(0.1,101))+
    scale_x_continuous(limits = c(-1, 50))+
    geom_hline(yintercept = 50)+
    geom_hline(yintercept = 10)
   
```

From this curves I read of the linear part of the response where we cross the 50% survival rate and use this than to calculate the linear models below.

###Resulting EC50 values for the strains tested under our drug concentrations

```{r}
#Here I use the filtered matrix to make the calculation of the EC50 more robust
#Calculate the model for WB
 modelWB <- lm(WB~Conc, data = MTZ_WB ,subset = Conc<11)

#summary(modelWB)
fWB <- round(coef(modelWB) ,10)

eqWB <- paste0("y = ", round(fWB[1],3),
             ifelse(sign(fWB[2])==1, " + ", " - "), abs(round(fWB[2],3)), " x ")

eWB <- round(solve(fWB[2], 50-fWB[1]) ,2)

EC50WB <- paste0("EC50[µM] = ", eWB[1])

gWB <- ggplot(MTZ_WB) + 
    ylab("Procent survival MTZ WB") + 
    geom_point (mapping = aes(x = Conc, y = WB), color="red")+
    theme_bw() +
    scale_x_continuous(limits = c(0, 15)) +
    scale_y_continuous(limits = c(0.1,105),trans = "log")+
    stat_smooth(aes(x = Conc, y = WB), method="lm", color="red")+
    annotate(geom="text", x=10, y=5, label=eqWB,
              color="black")+
    annotate(geom="text", x=10, y=2, label=EC50WB,
              color="black")

#Calculate the model for C6NR
 modelC6NR <- lm(C6NR~Conc, data = MTZ_C6NR ,subset = Conc<11)

#summary(modelC6NR)
fC6NR <- round(coef(modelC6NR) ,10)

eqC6NR <- paste0("y = ", round(fC6NR[1],3),
             ifelse(sign(fC6NR[2])==1, " + ", " - "), abs(round(fC6NR[2],3)), " x ")

eC6NR <- round(solve(fC6NR[2], 50-fC6NR[1]) ,2)

EC50C6NR <- paste0("EC50[µM] = ", eC6NR[1])

gC6NR <-ggplot(MTZ_C6NR) + 
    ylab("Procent survival MTZ C6NR") + 
    geom_point (mapping = aes(x = Conc, y = C6NR), color="blue")+
    theme_bw() +
    scale_x_continuous(limits = c(0, 15)) +
    scale_y_continuous(limits = c(0.1,105),trans = "log")+
    stat_smooth(aes(x = Conc, y = C6NR), method="lm", color="blue")+
    annotate(geom="text", x=10, y=5, label=eqC6NR,
              color="black")+
    annotate(geom="text", x=10, y=2, label=EC50C6NR,
              color="black")

#Calculate the model for C60.5
 modelC60.5 <- lm(C60.5~Conc, data = MTZ_C60.5 ,subset = Conc>4)

#summary(modelC60.5)
fC60.5 <- round(coef(modelC60.5) ,10)

eqC60.5 <- paste0("y = ", round(fC60.5[1],3),
             ifelse(sign(fC60.5[2])==1, " + ", " - "), abs(round(fC60.5[2],3)), " x ")

eC60.5 <- round(solve(fC60.5[2], 50-fC60.5[1]) ,2)

EC50C60.5 <- paste0("EC50[µM] = ", eC60.5[1])

gC60.5 <- ggplot(MTZ_C60.5) + 
    ylab("Procent survival MTZ C60.5") + 
    geom_point (mapping = aes(x = Conc, y = C60.5), color="green")+
    theme_bw() +
    scale_x_continuous(limits = c(10, 50)) +
    scale_y_continuous(limits = c(0.1,105),trans = "log")+
    stat_smooth(aes(x = Conc, y = C60.5), method="lm", color="green")+
    annotate(geom="text", x=30, y=5, label=eqC60.5,
              color="black")+
    annotate(geom="text", x=30, y=2, label=EC50C60.5,
              color="black")

#Calculate the model for C6aI
 modelC6aI <- lm(C6aI~Conc, data = MTZ_C6aI,subset = Conc>1 & Conc<50)

#summary(modelC6aI)
fC6aI <- round(coef(modelC6aI) ,10)

eqC6aI <- paste0("y = ", round(fC6aI[1],3),
             ifelse(sign(fC6aI[2])==1, " + ", " - "), abs(round(fC6aI[2],3)), " x ")

eC6aI <- round(solve(fC6aI[2], 50-fC6aI[1]) ,2)

EC50C6aI <- paste0("EC50[µM] = ", eC6aI[1])

gC6aI <- ggplot(MTZ_C6aI) + 
    ylab("Procent survival MTZ C6aI") + 
    geom_point (mapping = aes(x = Conc, y = C6aI), color="black")+
    theme_bw() +
    scale_x_continuous(limits = c(1, 50)) +
    scale_y_continuous(limits = c(0.1,105),trans = "log")+
    stat_smooth(aes(x = Conc, y = C6aI), method="lm", color="black")+
    annotate(geom="text", x=10, y=5, label=eqC6aI,
              color="black")+
    annotate(geom="text", x=10, y=2, label=EC50C6aI,
              color="black")

EC50valuesMTZ8_8_19 <- tribble(
   ~" ", ~"WB", ~"C6NR", ~"C60.5", ~"C6aI", 
   #--|--|----
  "EC50[µM]", eWB, eC6NR, eC60.5, eC6aI)

kable(EC50valuesMTZ8_8_19)
write_csv(EC50valuesMTZ8_8_19, path = "EC50MTZ8_19.csv")

vplayout <- function(x, y) viewport(layout.pos.row = x, layout.pos.col = y)
grid.newpage()
pushViewport(viewport(layout = grid.layout(2, 2)))
print(gWB, vp = vplayout(1, 1))
print(gC6NR, vp = vplayout(1, 2))
print(gC60.5, vp = vplayout(2, 1))
print(gC6aI, vp = vplayout(2, 2))
```
#Dicussion and conclusion


